---
title: "Global_presentation"
author: "Michał Michalak"
date: "21 01 2021"
output: ''
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r}
library(dplyr)
library(ggplot2)
library(ggpubr)
library(reshape2)
library(tibble)
library(sf)
library(tmap)
library(broom)
library(plotly)
library(magrittr)
```




#1. Read the input file

```{r}
world<-read.csv("world.csv", sep=";", dec=".", header = TRUE, check.names = FALSE)
```

#2. Assure that data on both infection and tests is supplied and that the number of tests is greater than that of infections

```{r}
world <-dplyr::filter(world, !is.na(total_tests) & !is.na(total_cases) & total_tests>total_cases )
```

```{r}
world <-dplyr::filter(world, 
                        #Asia peaks
                        date!='2020-06-24' & date!= '2020-08-06' &
                        
                        #Europe peaks
                        date!='2020-08-04' & date!= '2020-08-21' & date!= '2020-10-02' &
                        date!= '2020-07-04'& date!= '2020-12-02' & date!= '2020-12-14' &
                        date!= '2020-03-30' & date!= '2020-03-19'& date!= '2020-04-01' &
                        date!= '2020-03-27' & date!= '2020-03-18'& date!= '2020-04-12' &
                        
                        #Africa peaks
                        date!= '2020-04-07'&
                        date!= '2020-12-01'& 
                        
                        #North America peaks
                        date!= '2020-12-21' &
                        
                        #Oceania peaks
                        date!= '2020-08-22'& date!= '2020-08-28' & date!= '2020-08-29'&
                        date!= '2020-09-04'& date!= '2020-09-05' & date!= '2020-09-11'&
                        date!= '2020-09-12'& date!= '2020-09-18' & date!= '2020-09-19'&
                        date!= '2020-09-25'& date!= '2020-09-26' & date!= '2020-09-19'&
                        date!= '2020-10-03'& date!= '2020-10-04' & date!= '2020-10-05'&
                        date!= '2020-10-09'& date!= '2020-10-10' & date!= '2020-10-16'&
                        date!= '2020-10-17'& date!= '2020-10-23' & date!= '2020-10-24'&
                        date!= '2020-10-30'& date!= '2020-11-06' & date!= '2020-11-13'&
                        date!= '2020-11-14'& date!= '2020-11-20' & date!= '2020-11-21'&
                        date!= '2020-11-27'& date!= '2020-11-28' & date!= '2020-12-04'&
                        date!= '2020-12-11'& date!= '2020-07-28' & date!= '2020-07-31'&
                        date!= '2020-10-31'& date!= '2020-07-11' & date!= '2020-12-05'&
                        date!= '2020-12-18'& date!= '2020-03-22' & date!= '2020-03-29' &
                        date!= '2020-12-19'& date!= '2020-12-24' & date!= '2020-12-26' &
                      
                        #South America peaks
                        date!= '2020-04-20'& date!= '2020-05-29' & date!= '2020-06-30' &
                        date!= '2020-07-21'& date!= '2020-07-25' & date!= '2020-08-02' &
                        date!= '2020-08-08'& date!= '2020-08-15' & date!= '2020-08-31' &
                        date!= '2020-06-02'& date!= '2020-06-07' & date!= '2020-06-11' &
                        date!= '2020-06-14'& date!= '2020-06-15' & date!= '2020-06-16' &
                        date!= '2020-06-17'& date!= '2020-06-18' & date!= '2020-06-21'
                        )
```







#3. Select countries from Europe
#world <-dplyr::filter(world, continent=='Europe')

#4. Calculate local positivity

```{r}
world$LPR<-world$total_cases/world$total_tests`

```

#5. Subset the data frame with needed columns

```{r}
world <-dplyr::select(world, c("continent", "location", "date", "total_tests", "total_cases", "LPR", "population") )
```



#6. Sort the data frame by date rather than by countries

```{r}
world<- world[order(world$date),]
```



#7. Determine the time period

```{r}
date_range<-unique(world$date)
```



#8. Prepare empty vectors for global positivity, number of countries 
#for which data are available for each individual day, and
#global ratios needed for calculating the expected number of cases and tests

```{r}
global_ratio <- c()
number_countries <- c()

r_cases <- c()
r_tests <- c()
```

```{r}
for (k in date_range){ #for each date
  date_index <- which(world$date==k) #get row indices
  number_countries[k] <- length(date_index) #get number of countries 
  global_ratio[k] <- sum(  world$total_cases[date_index]    )/ sum(  world$total_tests[date_index]    ) #calculate global positivity
  r_cases[k] <-  sum(  world$total_cases[date_index]    )/sum(  world$population[date_index]    ) #calculate ratio for calculating the expected number of tests
  r_tests[k] <-  sum(  world$total_tests[date_index]    )/sum(  world$population[date_index]    ) #calculate ratio for calculating the expected number of cases
}
```




#9. Replicate global positivity with respect to the number of countries with a given date

```{r}
global_positivity <- rep(global_ratio, number_countries)
```



#10. Replicate ratios to calculated expected number of cases and tests with respect to the number of countries with a given date

```{r}
r_cases <- rep (r_cases, number_countries )
```

```{r}
r_tests <- rep (r_tests, number_countries )
```




#11. Assign global positivity as a new column of the data frame

```{r}
world$GPR <- global_positivity
```



#12. Assign WCSIR as a new column of the data frame

```{r}

world$WCSIR <-world$LPR/world$GPR
```


#13. Assign global ratio for calculating the expected number of cases as a new column of the data frame

```{r}
world$r_cases <- r_cases
```




#14. Assign global ratio for calculating the expected number of tests as a new column of the data frame

```{r}
world$r_tests <- r_tests
```



#15. Assign  expected number of cases as a new column of the data frame

```{r}
world$expected_cases <- world$r_cases*world$population
```



#16. Assign  expected number of tests as a new column of the data frame


```{r}
world$expected_tests <- world$r_tests*world$population
```



#17. Assign CSIR as a new column of the data frame
```{r}
world$CSIR <- world$total_cases/world$expected_cases
```




#18. Assign CSTR as a new column of the data frame
```{r}
world$CSTR <- world$total_tests/world$expected_tests
```


#19. Generate the plot of CSTR


```{r}
world$date <- as.Date(world$date)

```

```{r}
CSTR_plot <- ggplot(data=world[483:nrow(world),], aes(
  x=date,
  y=CSTR,
  group = location,
  colour = location
)
)+
  geom_line(size=0.5) +
  geom_point( size=1, shape=21, fill="white")+
  scale_x_date(labels = scales::date_format("%m-%Y"))+
  theme(axis.text.x=element_text(angle = -90, hjust = 0), legend.position = "none")
```

```{r}
CSTR_plot_inverse<- ggplot(data=world[483:nrow(world),], aes(
  x=date,
  y=1/CSTR,
  group = location,
  colour = location
)
)+
  geom_line(size=0.5) +
  geom_point( size=1, shape=21, fill="white")+
  scale_x_date(labels = scales::date_format("%m-%Y"))+
  theme(axis.text.x=element_text(angle = -90, hjust = 0), legend.position = "none")
```

```{r}
ggpubr::ggarrange(CSTR_plot,CSTR_plot_inverse, ncol =1, nrow=2, labels=c("a", "b"))
```




#20. Generate the plot of CSIR

```{r}
CSIR_plot <- ggplot(data=world[483:nrow(world),], aes(
  x=date,
  y=CSIR,
  group = location,
  colour = location
)
)+
  geom_line(size=0.5) +
  geom_point( size=1, shape=21, fill="white")+
  scale_x_date(labels = scales::date_format("%m-%Y"))+
  theme(axis.text.x=element_text(angle = -90, hjust = 0), legend.position = "none")
```



```{r}
CSIR_plot_inverse<- ggplot(data=world[483:nrow(world),], aes(
  x=date,
  y=1/CSIR,
  group = location,
  colour = location
)
)+
  geom_line(size=0.5) +
  geom_point( size=1, shape=21, fill="white")+
  scale_x_date(labels = scales::date_format("%m-%Y"))+
  theme(axis.text.x=element_text(angle = -90, hjust = 0), legend.position = "none")
```

```{r}
ggpubr::ggarrange(CSIR_plot,CSIR_plot_inverse, ncol =1, nrow=2, labels=c("a", "b"))
```


#21. Generate the plot of LPR


```{r}
LPR_plot <- ggplot(data=world[483:nrow(world),], aes(
  x=date,
  y=LPR,
  group = location,
  colour = location
)
)+
  geom_line(size=0.5) +
  geom_point( size=1, shape=21, fill="white")+
  geom_line( aes (x=date, y=GPR), size=0.5)+
  scale_x_date(labels = scales::date_format("%m-%Y"))+
  theme(axis.text.x=element_text(angle = -90, hjust = 0), legend.position = "none")
```


```{r}
LPR_plot_inverse<- ggplot(data=world[483:nrow(world),], aes(
  x=date,
  y=1/LPR,
  group = location,
  colour = location
)
)+
  geom_line(size=0.5) +
  geom_line(data=world[483:nrow(world),], aes (x=date, y=GPR), size=1.4)+
  geom_point( size=1, shape=21, fill="white")+
  scale_x_date(labels = scales::date_format("%m-%Y"))+
  theme(axis.text.x=element_text(angle = -90, hjust = 0), legend.position = "none")
```


```{r}
ggpubr::ggarrange(LPR_plot,LPR_plot_inverse, ncol =1, nrow=2, labels=c("a", "b"))
```


#22. Generate the plot of WCSIR

```{r}
WCSIR_plot <- ggplot(data=world[483:nrow(world),], aes(
  x=date,
  y=WCSIR,
  group = location,
  colour = location
)
)+
  geom_line(size=0.5) +
  geom_point( size=1, shape=21, fill="white")+
  scale_x_date(labels = scales::date_format("%m-%Y"))+
  theme(axis.text.x=element_text(angle = -90, hjust = 0), legend.position = "none")
```

```{r}
WCSIR_plot_inverse<- ggplot(data=world[483:nrow(world),], aes(
  x=date,
  y=1/WCSIR,
  group = location,
  colour = location
)
)+
  geom_line(size=0.5) +
  geom_point( size=1, shape=21, fill="white")+
  scale_x_date(labels = scales::date_format("%m-%Y"))+
  theme(axis.text.x=element_text(angle = -90, hjust = 0), legend.position = "none")
```


```{r}
ggpubr::ggarrange(WCSIR_plot,WCSIR_plot_inverse, ncol =1, nrow=2, labels=c("a", "b"))
```

#23. Save the resulting data frame
write.table(world, "world_modified_.csv", sep=";", dec=".", row.names=F)


